---
title: "R Notebook"
output: pdf_notebook
---

"Drug Consumption Quantified Dataset" 
The dataset contains information on the consumption of 18 central nervous system psychoactivedrugs. The dataset has 32 attributes and 1885 instances from respondents aged over 18 in English-speaking countries.

Objective:
We intend to evaluate an individual's risk of drug misuse based on personality characteristics such as Big Five personality traits (NEO-FFI-R), impulsivity (BIS-11),sensation seeking (ImpSS), and demographic information in English-speaking countries.

-Nonparametric Classification:
K-Nearest Neighbour
Random Forest
-Parametric Classification:
Logistic Regression
Naive Bayes
LDA/QDA

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(MASS) 
library(rpart)
library(rpart.plot)
library(tidyr)
library(corrplot)
library(ggplot2)
library(class)
library(e1071)
```

All input attributes are originally categorical and are quantified. After quantification values of all input features can be considered as real-valued.
Calculated polychoric correlation to quantify ordinal features.
Used Nonlinear Categorical Principal Component Analysis to quantify nominal features (gender, country, ethnicity).

```{r}
drug = read.csv("drug_consumption.csv", header = FALSE)
# remove index column
drug = drug[complete.cases(drug), -1]
names(drug) = c("Gender", "Age", "Education", "Country", "Ethnicity", "N", "E", "O", "A", "C", "Impulsive", "Sensation", "Alcohol", "Amphet", "Amyl", "Benzos", "Caff", "Cannabis", "Choc", "Coke", "Crack", "Ecstasy", "Heroin", "Ketamine", "Legalh", "LSD", "Meth", "Mushrooms", "Nicotine", "Semer", "VSA")

# year-based binary classification
# classify 'Never Used' + 'Used over a Decade Ago' + 'Used in Last Decade' as Non-user
changey = function(x) ifelse(x %in% c("CL0", "CL1", "CL2"), "NonUser", "User")
drug_y = drug %>% mutate_at(vars(13:31), changey) %>% mutate_if(is.character, as.factor)
levels(drug_y[13:31]) = c("Non-User", "User")

# decade-based binary classification
# classify 'Never Used' + 'Used over a Decade Ago' as Non-user
changed = function(x) ifelse(x %in% c("CL0", "CL1"), "NonUser", "User")
drug_d = drug %>% mutate_at(vars(13:31), changed)  %>% mutate_if(is.character, as.factor)
levels(drug_d[13:31]) = c("Non-User", "User")

# separate training and testing set
# can change drug_d to drug_y 
# knn/logistic/RF
set.seed(1)
n = nrow(drug)
train_id = sample(n, size = floor(n*0.7))
train = drug_d[train_id,]
train_x = train[, 1:12]
train_label = train[, 13:31]
test = drug_d[-train_id,]
test_x = test[, 1:12]
test_label = test[, 13:31]
# lda/qda/nb 
drug_U = which(drug_d$Alcohol == "User")
drug_NU = which(drug_d$Alcohol == "NonUser")
train_id = c(sample(drug_U, size = trunc(0.70 * length(drug_U))),
             sample(drug_NU, size = trunc(0.70 * length(drug_NU))))
# column index depends on classification for which drug (here alcohol-->col=13)
drug_train = drug_d[train_id, 1:13]
drug_test = drug_d[-train_id, 1:13]
```


```{r}
pca_data = as.matrix(train_x)
pca_results = princomp(pca_data, cor = F)
summary(pca_results)
plot(pca_results)
```
```{r EDA}
# example
str(train)
pairs(train[1:4], col=c("blue", "green")[train$Alcohol], 
      pch=c(1,2)[train$Alcohol])

dat.explore = gather(train[,1:13], key = "Variable", value = "Value", -c("Alcohol"))
ggplot(dat.explore) + geom_boxplot(aes(x = Alcohol, y = Value)) + facet_wrap(.~Variable, scales = "free_y") + theme_minimal()
```


